Analysis of gene regulatory properties underlying trait pleiotropy

### 2022-10-18 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH presentation

\begin{itemize}
\item Genetic loci with higher frequency in individuals with a given disease
\item A very active field with many studies
\item Many genetic variants are involved in several phenotypes
\end{itemize}
%
\vfill
%
Limitations
%
\begin{itemize}
\item Correlation between variants, ie. LD
\item Which cell types are causal to the disease?
\item Over 90\% GWAS variants fall in non-coding regions
\end{itemize}
%
\vfill
%
These limitations can be partially addressed by looking at eQTLs in different tissues

@@

\begin{itemize}
\item Colocalization analysis of 418 GWAS and 127 eQTL studies (53e3 pairs)
\item 143e3 shared causal variants with probability $\geq$ 0.8
\item Percentage of explained GWAS loci: 0$\%$ - 77$\%$ (Average 33$\%$)
\end{itemize}

%\begin{figure}[!]
%\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/poster_eccb22_barcelona/img/web.eps}
%\end{figure}

@@

Distance between GWAS traits based on eQTL beta

@@

Clustering coherent

@@

\small
\begin{itemize}
\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IL3, IL4, IL5, ...
\item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
\item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
\end{itemize}
\normalsize

@@

    \begin{center}
\includegraphics[width=0.8\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
     \end{center}
varianb
\let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Do pleiotropic variants associate to more egenes?}

@@

@@

>>>>

\begin{itemize}
\item Genetic alleles with higher frequency in individuals with a given disease
\item A very active field with many studies
\item Many genetic variants are involved in several phenotypes
\end{itemize}
%
\vfill
%
Limitations
%
\begin{itemize}
\item Correlation between variants
\item Which cell types are causal to the disease?
\item Over 90\% GWAS variants fall in non-coding regions
\end{itemize}
%
\vfill
%
These limitations can be partially addressed by looking at eQTLs in different tissues

@@

\begin{itemize}
\item Colocalization analysis of 418 GWAS and 127 eQTL studies (53e3 pairs)
\item 143e3 shared causal variants with probability $\geq$ 0.8
\item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
\end{itemize}

\begin{figure}[!]
\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/coloc_web.png}
\end{figure}

@@

Distance between GWAS traits based on eQTL/egene beta

@@

GWAS/eQTL colocalization analysis allows to cluster diseases coherently

@@

\small
\begin{itemize}
\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
\item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
\item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
\end{itemize}
\normalsize

@@

    \begin{center}
\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
     \end{center}

Upstream and downstream of genes and in the intron

\let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Why are these variants pleiotropic? Do pleiotropic variants associate to more egenes?}

@@

More egenes per variant-etissue pair.

@@

This could explain why pleiotropic variants active in more etissues and associated to more egenes
